---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stat545lamke07

<!-- badges: start -->
<!-- badges: end -->

The goal of `stat545lamke07` is to have a collection of functions that can 
quickly create toy data sets to test their statistical or machine learning
model on. Many times one is interested in using simulated data, but these
often need to be written out quickly. The `stat545lamke07` package therefore
aims to make this process easier and in an orderly way.

## Installation

You can install the released version of `stat545lamke07` from the 
[GitHub repository](https://github.com/lamke07/stat545lamke07)
with:

``` r
devtools::install_github("lamke07/stat545lamke07")
```

## Example

This is a basic example which shows you how to solve a common problem: the `generate_XY` function
creates a data set where $Y$ is a linear combination of the columns in $X$.
As such, a linear model on the full data set is expected to give a perfect fit.

```{r example}
library(stat545lamke07)

# Obtain a quick data set S = (X,Y)
df <- generate_XY()
print(head(df))

# Test a linear model
m1 <- lm(Y ~., data = df)
summary(m1)
```

It is possible to specify the individual parameters of the normal distribution for the columns of $X$:

* $n$: the desired number of data points in the data set. The corresponding parameter will be `n`.
* $\mu$: a $p$-dimensional vector of means. This can be any numeric vector. The corresponding parameter will be `mu`.
* $\sigma$: a $p$-dimensional vector of standard deviations. The values of $\sigma$ must be non-negative. The corresponding parameter will be `sigma`.
* $\beta$: a $p$-dimensional vector of coefficients. This can be any numeric vector. The corresponding parameter will be `beta_coefficients`.

Below we have given an example of how one could possibly specify the parameters. We need to make sure that all the dimensions are correct.

```{r example2}
# Obtain a quick data set S = (X,Y)
df <- generate_XY(n = 1000, mu = 1:10, sigma = 1:10, beta_coefficients = 1:10)
print(head(df))

# Test a linear model
m1 <- lm(Y~ X1 + X2 + X3 + X4, data = df)
summary(m1)
```
