---
title: "Causal Data Sets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Causal Data Sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup and Overview

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stat545lamke07)
```

The purpose of the `stat545lamke07` package is to be able to quickly test a method of interest
on a toy data set. To that end, the functions starting with `causal_` aim to 
create quick **causal** data sets where one can test a causal inference model.
In causal inference, one is usually interested in whether a treatment $T$ has a measurable impact
on the outcome $Y$. Often, there is associated data $X$ for the individuals.
Under the [Rubin Causal Model](https://en.wikipedia.org/wiki/Rubin_causal_model), we consider the outcomes $Y$ to have
potential outcomes $Y_0$ and $Y_1$, corresponding to the notion of "What if the patient had received treatment 0 (or treatment 1)?".
Under the *Fundamental Problem of Causal Inference*, both potential outcomes cannot be observed at the same time, which makes causal effect estimation hard.
The general idea is to use the `generate_XY()` function to generate the basic data set $S = (X,Y)$ and then add the treatment effects.

### Causal Data Sets with Binary Treatment

The function `causal_XTY_binary` builds on the `generate_XY()` function by selecting a treatment 1 with probability `treatment_prob`,
and treatment 0 otherwise. The `treatment_effect` is then added to $Y$ when treatment 1 is selected. Note that, however, we can
simulate the exact outcome $Y$ for both $Y_0$ and $Y_1$, since we have control over the generative probability model.

```{r}
df_causal <- causal_XTY_binary(n = 40, mu = 1:4, sigma = rep(2,4),
                  beta_coefficients = 1:4, treatment_prob = 0.75, treatment_effect = 25)

print(head(df_causal))
```

If we would like to turn this data set into a purely observational one, we only need to remove the potential outcomes.
This allows us to test causal inference techniques where the potential outcomes $Y_0, Y_1$ are unknown.

```{r}
df_causal %>%
  dplyr::select(-c("Y0", "Y1")) %>%
  head()
```

### Causal Data Sets with Multiple Treatments

Causal inference, however, need not be restricted to binary treatments only.
The function `causal_XTY_multiple` extends this notion of treatment to the case
where more than one treatment is available. Rather than one single treatment probability,
we now require a probability vector `treatment_prob` indicating how likely a treatment was chosen.
The treatment effect, `treatment_effect` will also be a vector that adds the overall treatment effect to
the existing outcomes $Y$, which serves as the baseline outcome. Again, we include the full information in 
the data set, and removing columns such as the potential outcomes $Y_i$ would turn this data set into
a causal data set to be analyzed.

```{r}
df_causal_multiple <- causal_XTY_multiple(n = 40, mu = rep(2, 5), sigma = 1:5,
                     beta_coefficients = 1:5,
                     treatment_prob = c(0.4, 0.1, 0.1, 0.2, 0.2),
                     treatment_effect = 1:5)

df_causal_multiple %>%
  head() %>%
  knitr::kable()
```



