---
title: "Normal Data Sets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Normal Data Sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup and Overview

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stat545lamke07)
```

The purpose of the `stat545lamke07` package is to be able to quickly test a method of interest
on a toy data set. To that end, the functions starting with `generate_` aim to 
create data sets based on the **normal** distribution.

### Generating only X

The key component of the `stat545lamke07` package is the `generate_X()` function,
which generates a data set $S = (X)$ where the columns of $X$ are normally distributed. The usage of such a data set is 
extremely flexible, as we can transform the data set quickly. To run the `generate_X()`,
all we requires is the number of data samples, as well as the right parametrization of $\mu$ and $\sigma$ (`mu` and `sigma`).

```{r}
df_X <- generate_X(n = 10, mu = rep(0,5), sigma = rep(2, 5))
print(head(df_X))
```

### Generating only X and Y

Suppose we would like to understand the effect of including more variables in our linear model.
In addition to just generating $X$ using `generate_X()`, we can now specify the exact linear coefficients
using the `beta_coefficients` parameter to obtain $$Y = X^T \beta$$ which leads to the data set
$S = (X,Y)$. Note that we need to ensure that the number of columns of $X$ are the same as the number of coefficients
in `beta_coefficients`. Ready to test this question with the use of `generate_XY()`, we first generate the data set.

```{r}
df <- generate_XY(n = 1000, mu = rep(0,10), sigma = rep(2,10), beta_coefficients = 1:10)
print(head(df))
```

Having generated the data set, we can now fit some linear models as well.

```{r}
# Test a linear model with 3 variables
m1 <- lm(Y~ X1 + X2 + X3, data = df)
summary(m1)

# Test a linear model with 6 variables
m2 <- lm(Y~ X1 + X2 + X3 + X4 + X5 + X6, data = df)
summary(m2)
```

We can quickly modify our assumptions about the data set by changing the relevant paramters in the `generate_XY` function,
namely `mu`, `sigma`, and `beta_coefficients`.

```{r}
df <- generate_XY(n = 1000, mu = 51:55, sigma = seq(10,15, length.out = 5), beta_coefficients = 21:25)
print(head(df))
```

### Generating X with categorical variables

So far we have assumed that $X$ contains only continuous variables. However, it is also possible to include 
categorical variables in the data set. To this end, we have written the `generate_X_cat()` function that 
additionally computes categorical factors, which can be achieved through the `no_of_cat` parameter.
`no_of_cat = c(4,5)` is a vector where each entry corresponds to the number of categories in each column.
In this case, we would have one column with 4 categories and 5 categories each.

```{r}
df_cat <- generate_X_cat(n = 40, mu = 1:5, sigma = rep(1, 5), no_of_cat = c(2,3))
print(head(df_cat))
```




